---
title: "COVID-19 in South Africa"
description: This article mainly discusses the COVID-19 situation and affection to South Africa in 2021 (until Aug 29, 2021).
author:
  - name: Yunqi Chen
    url: https://janieyunqichen.netlify.app/
date: "`r Sys.Date()`"
output: distill::distill_article
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align="center")
library(tidyverse)
library(gridExtra)
library(lubridate)
library(kableExtra)
library(sparkline)
library(DT)
library(gt)
```

# Data Description

This article discusses the COVID-19 situation and affection to South Africa in 2021 (until Aug 29, 2021). The data in this article is downloaded from *Our World in Data*^[Hannah Ritchie, Edouard Mathieu, Lucas Rodés-Guirao, Cameron Appel, Charlie Giattino, Esteban Ortiz-Ospina, Joe Hasell, Bobbie Macdonald, Diana Beltekian and Max Roser (2020) - "Coronavirus Pandemic (COVID-19)". Published online at [OurWorldInData.org](http://ourworldindata.org/). Retrieved from: '[https://ourworldindata.org/coronavirus](https://ourworldindata.org/coronavirus)' [Online Resource]].

This article uses the data of South Africa from Jan1 2021, to Aug 29, 2021. The data for the first section of *The Fact of COVID-19 Cases in South Africa* contains the date, total cases, new cases, total deaths, and new deaths. The second section of *Nucleic Acid Testing for COVID-19 Situation* data includes the date, total test number per thousand people, new test number per thousand people, and positive rate. The third section of *Covid-19 Vaccination Status* data contains the date, population, and total vaccinations number.

```{r}
#Tidy process
covidcase_raw <- read_csv(here::here("data", "owid-covid-data.csv")) %>% 
  filter(location == "South Africa",
         date > as.Date('2021-01-01'))

case <- covidcase_raw %>% 
  select(date, total_cases, new_cases, total_deaths,new_deaths)

test <- covidcase_raw %>% 
  select(date, total_tests_per_thousand, new_tests_per_thousand, positive_rate)


vaccine <- covidcase_raw %>% 
  select(date, total_vaccinations, population)
```

# The Fact of COVID-19 Cases in South Africa

Until Aug 29, total COVID-19 cases in South Africa have reached over 2.6 million, and deaths cases over 77 thousand. From table \@ref(fig:t1), the new COVID-19 cases trend is generally a straight line every month. This indicates there are almost average new cases every day. The average number of new deaths cases is very different each month. By sorting the New Deaths/day column, January 2021 has the most deaths cases with a 509 per day averagely, and April 2021 has the least with 50 deaths per day. From the New Deaths Trend/month, except for February, March, and June, the other months fluctuate around the average deaths case number. February, March, and June have an unusual peek at someday.

```{r t1, fig.cap="In 2021, new COVID-19 cases in South Africa climb at a constant rate every day from 1.3 to 2.6 million. New death cases generally fluctuate around the average each month."}
case_month <- case %>% 
  mutate(month = month(date)) %>% 
  select(-date)

case_month %>% 
  group_by(month) %>% 
  summarise(avg_total_cases = scales::comma(round(mean(total_cases))),
            avg_new_cases = scales::comma(round(mean(new_cases))),
            new_cases_trend = spk_chr(total_cases, type = "line"),
            avg_total_deaths = scales::comma(round(mean(total_deaths))),
            avg_new_deaths = round(mean(new_deaths)),
            new_deaths_trend = spk_chr(new_deaths, type = "line")) %>% 
  select(-month) %>% 
  datatable(escape = FALSE, 
            options = list(columnDefs = list(list(className = 'dt-right', targets = 0:4))),
            colnames = c("Month", "Total Cases", "New Cases/day", "New Cases Trend/month","Total Deaths", "New Deaths/day", "New Deaths Trend/month")
            ) %>%
  spk_add_deps()
```

# Nucleic Acid Testing for COVID-19 Situation

As the table \@ref(fig:t2) indicates, the positive rate was less than 10% from February to May 2021. However, the new test number of people is around 0.4-0.5 thousand people per day, which is almost half of the other months in 2021. In late August, there are more people have been tested than ever in 2021.

```{r t2, fig.cap="The positive rate was comparably low from February to May. This may be because there were fewer people took vaccine test during that time."}
test_month <- test %>% 
  mutate(month = month(date)) %>% 
  select(-date) %>% 
  na.omit()

test_month %>% 
  group_by(month) %>% 
  summarise(avg_positive_rate = scales::label_percent()(mean(positive_rate)),
            avg_total_tests_per_thousand = round(mean(total_tests_per_thousand)),
            avg_new_tests_per_thousand = round(mean(new_tests_per_thousand),2),
            tests_per_thousand_trend = spk_chr(new_tests_per_thousand, type = "line"),
            avg_positive_rate = scales::label_percent()(mean(positive_rate))
           ) %>% 
  select(-month) %>% 
  datatable(escape = FALSE, 
            options = list(columnDefs = list(list(className = 'dt-right', targets = 0:4))),
            colnames = c("Month", "Positive Rate(%)", "Total Test (1000 ppl)", "New Test (1000 ppl/day)", "New Test Trend/month")
            ) %>%
  spk_add_deps()
```

# Covid-19 Vaccination Status

February 2021, the COVID-19 vaccine has entered South Africa. Since June, the vaccination rate was increasing at a remarkable rate. Until Aug 30, there are almost 20% with 11 million people have been vaccinated.

```{r}
vaccine_month <- vaccine %>% 
  na.omit() %>% 
  mutate(month = month(date)) %>% 
  select(-date)

vt <- vaccine_month %>% 
  group_by(month) %>% 
  summarise(max_total_vaccinations = round((max(total_vaccinations)))
            ) %>% 
  mutate(rate = max_total_vaccinations/60041996)
  
vt$max_total_vaccinations <- scales::comma(vt$max_total_vaccinations)
vt$rate <- scales::label_percent()(vt$rate)

gt(vt) %>% 
  tab_header(title = "Vaccination Status in South Africa",
             subtitle = "Cumulative number and percentage of COVID-19 vaccination in South Africa from February to August.") %>% 
  tab_source_note(md("Source: Our World in Data, accessed at Aug29 2021")) %>% 
  cols_label(month = "Month",
             max_total_vaccinations = "Number of People Vaccinated",
             rate = "Percentage of Vaccinated People(%)") %>% 
  cols_align(
  align = "right",
  columns = max_total_vaccinations) %>% 
  cols_align(
  align = "right",
  columns = rate)
```






# Reference

## R Packages
- Baptiste Auguie (2017). gridExtra: Miscellaneous Functions for "Grid" Graphics. R
  package version 2.3. https://CRAN.R-project.org/package=gridExtra

- Garrett Grolemund, Hadley Wickham (2011). Dates and Times Made Easy with
  lubridate. Journal of Statistical Software, 40(3), 1-25. URL
  https://www.jstatsoft.org/v40/i03/.
  
- Hao Zhu (2021). kableExtra: Construct Complex Table with 'kable' and Pipe Syntax.
  R package version 1.3.4. https://CRAN.R-project.org/package=kableExtra

- Ramnath Vaidyanathan, Kent Russell and Gareth Watts (2016). sparkline: 'jQuery' Sparkline
  'htmlwidget'. R package version 2.0.

- R Core Team (2020). R: A language and environment for statistical computing. R
  Foundation for Statistical Computing, Vienna, Austria. URL
  https://www.R-project.org/.
  
- Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source
  Software, 4(43), 1686, https://doi.org/10.21105/joss.01686

- Yihui Xie, Joe Cheng and Xianying Tan (2021). DT: A Wrapper of the JavaScript Library
  'DataTables'. R package version 0.18. https://CRAN.R-project.org/package=DT

## Websites

- Hannah Ritchie, Edouard Mathieu, Lucas Rodés-Guirao, Cameron Appel, Charlie Giattino, Esteban Ortiz-Ospina, Joe Hasell, Bobbie Macdonald, Diana Beltekian and Max Roser (2020) - "Coronavirus Pandemic (COVID-19)". Published online at OurWorldInData.org. Retrieved from: 'https://ourworldindata.org/coronavirus' [Online Resource]
